---
title: "outline of questions"
author: "Wei Wang"
date: 2017-02-07
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`


## list of questions

let's write the $X$ into $2 \times2$ partitions.

$$X=  \begin{pmatrix} X_{11} & X_{12} \\ X_{21} & X_{22}  \end{pmatrix}= \left[ \begin{array}{c} U_{1} \\ U_{2} \end{array} \right] \Gamma \left[ \begin{array}{cc} V_{1}^{T}& V_{2}^{T} \end{array} \right] $$

### 1 algorithm to find original svd

we want to find a solution for 

\begin{eqnarray}
&\min_{U,V,\Gamma}& ||X_{11} - U_1 \Gamma V_1^T||_2 + ||X_{22} - U_2 \Gamma V_2^T||_2 \\
&st& U_1 U_1^T +  U_2 U_2^T = I \\
& & V_1 V_1^T +  V_2 V_2^T = I
\end{eqnarray}

#### 1.1 convexity of this problem

- We are not sure if this optimazation problem is convex or not.

#### 1.2 uniqueness of the solution

- The solution of the above problem is unique or not?

#### 1.3 computational complexity

- To solve this problem might be more computationaly intensive comparing with BCV which just need to calculate svd and general inverse.

### 2 separate svd vs original svd

we know that 

\begin{eqnarray}
X_{11} = U_1 \Gamma V_1^T \\
X_{22} = U_2 \Gamma V_2^T
\end{eqnarray}

but $U_1 \Gamma V_1^T$ is not SVD for $X_{11}$. On the other hand, if we combine the svd for $X_{11} = \tilde{U}_1 \Sigma_{11}\tilde{V}_1$ and that for $X_{22} = \tilde{U}_2 \Sigma_{22}\tilde{V}_2$, the combining columns $\left[ \begin{array}{c} \tilde{U}_{1} \\ \tilde{U}_{2} \end{array}  \right]$ is not the original svd $\left[ \begin{array}{c} U_{1} \\ U_{2} \end{array} \right]$. It's because $\tilde{U}_{1}$ and $\tilde{U}_{2}$ just contains local information for $X_{11}$ and $X_{22}$.

