---
title: "Initialization for missing data"
author: "Wei Wang"
date: 09-16
output: 
  html_document:
    code_folding: hide
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

```{r}
sim_hd = function(N, P, SF, SL, signal, a = rchisq(N,3),b = rchisq(P,1),mu = 0){
  
  E = matrix(rep(0,N*P),nrow=N)
  sig2_true = matrix(rep(0,N*P),nrow=N)
  for(i in 1:N){
    for(j in 1:P){
      sig2_true[i,j] = mu + a[i] * b[j]
      E[i,j] = rnorm(1,0,sqrt(mu + a[i] * b[j]))
    }
  }
  
  K=1
  lstart = rnorm(N, 0, signal)
  
  fstart = rnorm(P, 0, signal)
  
  index = sample(seq(1:N),(N*SL))
  lstart[index] = 0
  index = sample(seq(1:P),(P*SF))
  fstart[index] = 0
  
  Y = lstart %*% t(fstart) + E
  
  return(list(Y = Y, L_true = lstart, F_true = fstart, Error = E,sig2_true = sig2_true))
}
sim_K = function(K, N, P, SF, SL, signal,noise){
  E = matrix(rnorm(N*P,0,noise),nrow=N)
  Y = E
  L_true = array(0, dim = c(N,K))
  F_true = array(0, dim = c(P,K))
  
  for(k in 1:K){
    lstart = rnorm(N, 0, signal)
    fstart = rnorm(P, 0, signal)
    
    index = sample(seq(1:N),(N*SL))
    lstart[index] = 0
    index = sample(seq(1:P),(P*SF))
    fstart[index] = 0
    
    L_true[,k] = lstart
    F_true[,k] = fstart
    
    Y = Y + lstart %*% t(fstart)
  }
  return(list(Y = Y, L_true = L_true, F_true = F_true, Error = E))
}
```


```{r}
library(flashr2)
```

## same data set as in "flash_beta_test"

```{r,message=F,warning=F}
library(softImpute)
set.seed(999)
N = 100
P = 200
dat = sim_K(K=1,N, P, SF = 0.5, SL = 0.5, signal = 3,noise = 5)
# a = rchisq(N,3)
# b = rchisq(P,1)
# data = sim_hd(N, P, SF = 0.5, SL = 0.5, signal = 1, a ,b,mu = 0)
Y = dat$Y
Y_miss = Y
Y_miss[1:50,1:100] = NA
Y_miss[51:100,101:200] = NA
```

Here I just run flash on this data sets four times and provide different result due to the unstable initial value.

```{r}
par(mfrow = c(2,2))
for(i in 1:4){
  data.miss = set_flash_data(Y_miss)
  g1.miss = flash_r1(data.miss,verbose=F)
  plot(as.vector(dat$L_true %*% t(dat$F_true)), as.vector(g1.miss$EL %*% t(g1.miss$EF)), xlab = "truth",ylab = "estimated", main = "vectorized LF")
}
```


## I don't think softimpute is good initialization method (or maybe I don't understand this method very well)

Here is a example of use softimpute. maybe I was not using it in the right way.

Here I use default `lambda`, and we can try `type = "svd"`. I tried different combination and this method is still not stable.

```{r}
par(mfrow = c(2,2))
for(i in 1:4){
  gsoft = softImpute(Y_miss, rank.max = 1,type = "als")
  sfl = (gsoft$u)
  sff = gsoft$d * (gsoft$v)
  plot(as.vector(dat$L_true %*% t(dat$F_true)), as.vector(sfl %*% t(sff)), xlab = "truth",ylab = "estimated", main = "vectorized LF")
}
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
